### 2019-12-17 Papers/Torque_Abstract/main.tex
small corrections to the Torque abstract

\begin{addmargin}[2.5cm]{0em}% 1em left, 2em right
\textbf{P. Kumar$^{\boldsymbol{\mathsf{1}}}$, B. Sanderse$^{\boldsymbol{\mathsf{1}}}$}

$^{\mathsf{1}}$ Centrum Wiskunde \& Informatica, Science Park 123, Amsterdam.

@@

% =============================================
\section{Introduction}
Aeroelastic models such as the Blade Element Momentum (BEM) method \cite{HandBook} continue to play a critical role in the design, development and optimization of modern wind turbines. The accuracy of BEM predictions is affected by uncertainties and inaccuracies, for e.g., in the external conditions (wind parameters), in the turbine specifications (geometric parameters), and in the BEM equations itself (model parameters). For the purpose of design, uncertainty quantification and optimization, it is crucial to limit the number of parameters, typically by performing sensitivity studies. Most studies focused on the effect of uncertainties in the external conditions and geometry, see e.g. \cite{Echeverria2017,Robertson2018}. % removed reference Matthaus2017,Murcia2018 
\section{Objectives}
The long-term objective of this study is to develop calibrated BEM models that give users an indication of the uncertainty associated with the predictions (loads, power, etc.) originating not only from external conditions and geometry but also from the model formulation itself. For this purpose, we will calibrate the model parameters present in BEM models. Examples of such model parameters are the time constant in dynamic stall models, the wake correction factor, the tip loss model parameter, and the lift- and drag-polars \cite{Sayed2019}. To limit the number of model parameters involved in the calibration process, \textit{the objective of the current study is to perform a global sensitivity study of the outputs of the BEM model towards both geometric and model uncertainties}.

@@

% =============================================
\section{Methodology}
To compute parameter sensitivities we use a global sensitivity analysis (GSA) based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Rinker2016}. We employ the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}, which computes the Sobol indices from a sparse polynomial chaos expansion. \texttt{UQLab}'s modular structure allows for easy integration with available BEM codes.

@@

% =============================================
\section{Results}
The aeroelastic code that we use is the \texttt{ECN Aero-Module} \cite{Boorsma2012} and the turbine is the 2MW NM80 turbine (blade radius of 38.8m) from the DANAERO project \cite{Troldborg2013}. The data for airfoil lift- and drag-polars are available at four locations along the blade at 11.87m, 17.82m, 28.97m and 35.53m. %The reference value of polars are obtained from wind-tunnel experiment with 3D corrections. 
The lift and drag variables at these sections are numbered Cl1 - Cl4 and Cd1 - Cd4 respectively. Uncertainty enters by multiplying the reference lift and drag curves by a random variable (see Fig. \ref{fig:samples}). The chord and twist are perturbed via a number of control points that define the NURBS representation; we use 5 chord (Ch1 - Ch5) and 6 twist (Tw1 - Tw6) control points and perturb each value randomly following a uniformly distribution. Examples of random samples for these input parameters are presented in Fig.~\ref{fig:samples}.

>>>>

\begin{addmargin}[2.5cm]{0em}% 1em left, 2em right
\textbf{P. Kumar$^{\boldsymbol{\mathsf{1}}}$, B. Sanderse$^{\boldsymbol{\mathsf{1}}}$, K. Boorsma$^{\boldsymbol{\mathsf{2}}}$, M. Caboni$^{\boldsymbol{\mathsf{2}}}$}

$^{\mathsf{1}}$ Centrum Wiskunde \& Informatica, Science Park 123, Amsterdam, The Netherlands.
$^{\mathsf{2}}$ TNO, Westerduinweg 3, Petten, The Netherlands.

@@

% =============================================
\section{Introduction}
Aeroelastic models such as the Blade Element Momentum (BEM) method \cite{HandBook} continue to play a critical role in the design, development and optimization of modern wind turbines. The accuracy of BEM predictions is affected by uncertainties and inaccuracies, for e.g., in the external conditions (wind parameters), in the turbine specifications (geometric parameters), and in the BEM equations itself (model parameters). For the purpose of design, uncertainty quantification and optimization, it is crucial to limit the number of parameters, typically by performing sensitivity studies. Most studies so far focused on the effect of uncertainties in the external conditions and in geometry, see e.g.~\cite{Echeverria2017,Robertson2018}. % removed reference Matthaus2017,Murcia2018 
\section{Objectives}
The long-term objective of this study is to develop calibrated BEM models that give users an indication of the uncertainty associated with the predictions (loads, power, etc.) originating not only from external conditions and geometry but also from the model formulation itself. For this purpose, we will calibrate the model parameters present in BEM models. Examples of such model parameters are the time constant in dynamic stall models, the wake correction factor, the tip loss model parameter, and the lift- and drag-polars \cite{Sayed2019}. To limit the number of model parameters involved in the calibration process, \textit{the objective of the current study is to perform a global sensitivity study of the outputs of the BEM model towards both geometric and model uncertainties}.

@@

% =============================================
\section{Methodology}
To compute parameter sensitivities we use a global sensitivity analysis (GSA) based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Rinker2016a}. We employ the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}, which computes the Sobol indices from a sparse polynomial chaos expansion. \texttt{UQLab}'s modular structure allows for easy integration with available BEM codes.

@@

% =============================================
\section{Results}
The aeroelastic code that we use is the BEM model of the \texttt{ECN Aero-Module} \cite{Boorsma2012} and the turbine is the 2MW NM80 turbine (blade radius of 38.8m) from the DANAERO project \cite{Madsen2010a}. The data for the airfoil lift- and drag-polars is available at four locations along the blade: at 11.87m, 17.82m, 28.97m and 35.53m. %The reference value of polars are obtained from wind-tunnel experiment with 3D corrections. 
The lift and drag variables at these sections are numbered Cl1 - Cl4 and Cd1 - Cd4 respectively. Uncertainty enters by multiplying the reference lift and drag curves by a random variable (see Fig. \ref{fig:samples}). The chord and twist are perturbed via a number of control points that define the NURBS representation; we use 5 chord (Ch1 - Ch5) and 6 twist (Tw1 - Tw6) control points and perturb each value randomly following a uniform distribution. Examples of random samples for these input parameters are presented in Fig.~\ref{fig:samples}.
